---
title: "Homework 2"
author: "Aakriti Gupta"
date: 2018-10-03
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```


## **Problem 1**
### *Import and clean the CSV*
 
```{r}
nyc_transit = read_csv(file = "./data/NYC_Transit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE)) 
```

This dataset contains information about the NYC subway stations, lines and routes served by them and whetehr they have options for vending or not or whether tehy are ADA compliant or not.

I imported the csv using thr readr package and then cleaned using janitor package. Then, I used the dplyr package to select certain variables and mutate the entry variable from a character to a logical variable.

There are 465 unique stations:

```{r}
 nyc_transit %>% 
  distinct(line, station_name) 
```

The number of stations that are ADA compliant are 468 
```{r}
count(nyc_transit, ada)
```

Reformatting the dataset:

```{r}
nyc_transit_tidy = nyc_transit %>%
  gather(key = "route_number", value = "route_name", route1:route11) %>%
  select(route_name, route_number, everything()) %>% 
  group_by(route_name, route_number) %>%
  separate(route_number, into = c("route_str", "number"), sep = 5) %>% 
  select(-route_str) 
```

The A train serves 60 distinct stations:

```{r}
nyc_transit_tidy %>%
  group_by(route_name) %>%
  filter(route_name == "A") %>% 
  distinct(route_name, line, station_name)
```


## **Problem 2**
### *Importing and cleaning the trash wheel sheet*

```{r}
trash_wheel = readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, range = "A2:N258") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(
    sports_balls = NULL, 
    sports_balls_i = as.integer(sports_balls))
```


### *Importing and cleaning the sheet for 2016 precipitation data*
```{r}
ppt_2016 = readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4, skip = 1) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>% 
  mutate(year = 2016)
```

### *Importing and cleaning the sheet for 2017 precipitation data*
```{r}
ppt_2017 = readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3, skip = 1) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total), !is.na(month)) %>% 
  mutate(year = 2017)
```

### *Binding the rows to have a combined dataset and changing the month column from numbers to names*

```{r}
ppt_tidy = bind_rows(ppt_2016, ppt_2017) %>%
  janitor::clean_names() %>%
  select(year, everything()) %>% 
  mutate(month =  month.name[month])
```

Mr. Trash Wheel is “a water-wheel vessel that removes trash from the Inner Harbor in Baltimore, Maryland.” It (or he) sits at an intake into the Inner Harbor and intercepts litter and debris carried by the Jones Falls River toward the harbor. It has removed over a million pounds of litter since May 2014!
The total precipitation in 2017 was `r ppt_tidy %>% filter(year == 2017) %>% summarize(total_ppt_2017= sum(total))`. The median number of sportsballs in a dumpster in 2016 was `r trash_wheel %>% filter(year == 2016) %>% summarize(median_sportsballs_2016= median(sports_balls_i, na.rm = TRUE))`


###Problem 3

Loading the dataset

```{r}
library(p8105.datasets)
data(brfss_smart2010)
```

Cleaning, wrangling and tidying data:

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>%
  mutate(response = tolower(response)) %>% 
  spread(key = "response", value = "data_value") %>% 
  janitor::clean_names() %>% 
  mutate(excellent_or_very_good = excellent + very_good)
``` 

(i) Unique locations represented in the dataset are:

```{r}
brfss %>% 
  group_by(locationabbr) %>% summarise(count = n()) 
```
 
 This yields 51 unique locations. Every state is represented. New Jersey is represented the most with frequency of 146.
 
 (ii)
 
```{r}
brfss %>% 
  group_by(year) %>% 
  filter(year == 2002) %>% 
  summarize(median = median(excellent, na.rm = TRUE))
```
 For the year 2002, the median value for excellent response is 23.6.
 
 (iii) Histogram
 
```{r}
brfss %>% 
  group_by(year) %>% 
  filter(year == 2002) %>% 
ggplot(aes(x = excellent)) + geom_histogram()
```
 
 (iv) Scatterplot
```{r}
 brfss %>% 
 filter(locationabbr == "NY") %>% 
  separate(locationdesc, into = c("state", "county"), sep = "-") %>%
  select(year, county, excellent) %>%
  spread(key = "county", value = "excellent") %>%
  janitor::clean_names() %>%
  select(year, queens_county, new_york_county) %>%
  ggplot(aes(x = new_york_county, y = queens_county)) + 
  geom_point(aes(color = year))
```
 