---
title: "Homework 2"
author: "Aakriti Gupta"
date: 2018-10-03
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## **Problem 1**
### *Import and clean the CSV*
 
```{r}
nyc_transit = read_csv(file = "./data/NYC_Transit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE)) 
  
```

Write a short paragraph about the dataset above.

(i) There are 465 unique stations:

```{r}
 nyc_transit %>% 
  distinct(line, station_name) %>% View 
```

(ii) The number of stations that are ADA compliant are 468 `r count(nyc_transit, ada)`
 
 
(iii) The number of entrances/exits without vending that allow entry are 

Reformatting the dataset:

```{r}
nyc_transit_tidy = nyc_transit %>%
  gather(key = "route_number", value = "route_name", route1:route11) %>%
  select(route_name, route_number, everything()) %>% 
  group_by(route_name, route_number) %>%
  separate(route_number, into = c("route_str", "number"), sep = 5) %>% 
  select(-route_str) %>% 
View
```


The A train serves 60 distinct stations:

```{r}
nyc_transit_tidy %>%
  group_by(route_name) %>%
  filter(route_name == "A") %>% 
  distinct(route_name, line, station_name) %>% View
```



 
## **Problem 2**
### *Importing and cleaning the trash wheel sheet*

```{r}
trash_wheel = readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, range = "A2:N258") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(
    sports_balls = NULL, 
    sports_balls_i = as.integer(sports_balls))
```

Median number of sportsballs in a dumpster in 2016 was`r trash_wheel %>% filter(year == 2016) %>% summarize(median_sportsballs_2016= median(sports_balls_i, na.rm = TRUE))`

### *Importing and cleaning the sheet for 2016 precipitation data*
```{r}
ppt_2016 = readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4, skip = 1) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>% 
  mutate(year = 2016)
```
 
### *Importing and cleaning the sheet for 2017 precipitation data*
```{r}
ppt_2017 = readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3, skip = 1) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total), !is.na(month)) %>% 
  mutate(year = 2017)
```

### *Binding the rows to have a combined dataset and changing the month column from numbers to names*

```{r}
ppt_tidy = bind_rows(ppt_2016, ppt_2017) %>%
  janitor::clean_names() %>%
  select(year, everything()) %>% 
  mutate(month =  month.name[month])
```


 `r sum(ppt_tidy$total)`
 `r ppt_tidy %>% filter(year == 2017) %>% summarize(total_ppt_2017= sum(total))`
 
###Problem 3

Loading the dataset

```{r}
library(p8105.datasets)
data(brfss_smart2010)
```

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>%
  mutate(response = tolower(response)) %>% 
  spread(key = "response", value = "data_value") %>% 
  janitor::clean_names() %>% 
  mutate(excellent_or_very_good = excellent + very_good)
``` 

 
 (i) Unique locations represented in the dataset are:

```{r}
brfss %>% 
  group_by(locationabbr) %>% summarise(count = n()) %>% View
```
 
 This yields 51 unique locations. Every state is represented. New Jersey is represented the most with frequency of 146.
 
 (ii)
 
```{r}
brfss %>% 
  group_by(year) %>% 
  filter(year == 2002) %>% 
  summarize(median = median(excellent, na.rm = TRUE))
```
 For the year 2002, the median value for excellent response is 23.6.
 
 (iii) Histogram
 
```{r}
brfss %>% 
  group_by(year) %>% 
  filter(year == 2002) %>% 
ggplot(aes(x = excellent)) + geom_histogram()
```
 
 (iv) Scatterplot
```{r}
 brfss %>% 
 filter(locationabbr == "NY") %>% 
  separate(locationdesc, into = c("state", "county"), sep = "-") %>%
  select(year, county, excellent) %>%
  spread(key = "county", value = "excellent") %>%
  janitor::clean_names() %>%
  select(year, queens_county, new_york_county) %>%
  ggplot(aes(x = new_york_county, y = queens_county)) + 
  geom_point(aes(color = year))
```
 

 
 